### 2021-10-02 metahit/paper_analyses.Rmd
Fixed hyperparameters / empirical Bayes rework

---
output: html_document
editor_options: 
  chunk_output_type: console
---
Keep this nice and tidy.

```{r}

setwd("~/work/chronic/disbayes/metahit")
source("paper_analyses_header.r")
source("read_model_results.r")

```

# Data summary

Raw point estimates of mort, inc and prev.   Just women for now.  Curves compared between areas, faceted by disease.   Shows diseases where prevalence and incidence are very low.  Mild differences between areas.

```{r,eval=FALSE}

foo <- function(x){
  cat(deparse(substitute(x)))
}

ht <- 3
pdf("../../write/mort_data.pdf", height=ht)
gbdplot$areagender <- interaction(gbdplot$area, gbdplot$gender)
ggplot(gbdplot, aes(x=age, y=mort_prob, col=area, lty=gender, group=areagender)) + 
  facet_wrap(~disease, ncol=3, scales="free_y") + 
  coord_cartesian(xlim = c(50,100)) +
  ylab("Annual mortality probability") + 
  xlab("Year of age") +
  geom_ribbon(aes(ymin=mort_lower, ymax=mort_upper), fill="gray80", col="gray80") +
  geom_line() + 
  labs(col = "City region") +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 7.5),
        legend.position = "none")
dev.off()

pdf("../../write/inc_data.pdf", height=ht)
ggplot(gbdplot, aes(x=age, y=inc_prob, col=area, lty=gender, group=areagender)) + 
  facet_wrap(~disease, ncol=3, scales="free_y") + 
  coord_cartesian(xlim = c(50,100)) +
  ylab("Annual incidence probability") + 
  xlab("Year of age") +
  geom_ribbon(aes(ymin=inc_lower, ymax=inc_upper), fill="gray80", col="gray80") +
  geom_line() + 
  labs(col = "City region") +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 7.5),
        legend.position = "none")
dev.off()

pdf("../../write/prev_data.pdf", height=ht)
ggplot(gbdplot, aes(x=age, y=prev_prob, col=area, lty=gender, group=areagender)) + 
  facet_wrap(~disease, ncol=3, scales="free_y") + 
  coord_cartesian(xlim = c(50,100)) +
  ylab("Prevalence probability") + 
  xlab("Year of age") +
  geom_ribbon(aes(ymin=prev_lower, ymax=prev_upper), fill="gray80", col="gray80") +
  geom_line() + 
  labs(col = "City region", lty="") +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 7.5),
        legend.spacing.y = unit(0.1, "cm"))
dev.off()

```

# Independent areas

* Results from fitting disbayes model by disease, gender and area separately

* Stomach cancer and uterine cancer are done using national data.

```{r,eval=FALSE}

## Convergence checks.   All OK with 5000 samples and the occasional model simplification
resnc <- resnh %>% filter(Rhat > 1.05)
table(resnc$disease, resnc$area, resnc$gender)

## Compare modal approx with posterior 
## Page of plots for a single area and gender.  Disease variation more important than area/gender variation  
## stmc, utrc use national point estimates 
## What are the current settings used for these?

resall %>% 
  filter(var=="cf", disease=="Dementia", gender=="Female", 
         area=="Bristol", between(age, 70, 90), model=="Independent areas")

resoptunc <- readRDS("resoptunc_nonhier.rds") %>% 
  filter(!disease %in% c("Stomach cancer","Uterine cancer"))
resoptnat <- readRDS("resopt_nat.rds")
resoptunccf <- resoptunc %>%
  full_join(resoptnat) %>% 
  filter(var=="cf") %>%
  mutate(p50 = 1 - exp(-`par`),
         plow = 1 - exp(-`2.5%`),
         pupp = 1 - exp(-`97.5%`)) %>%
  filter(area=="Bristol" | area=="England", gender=="Female") %>%
  mutate(disease = fct_recode(disease, !!!disease_shorten),
           disease = factor(disease, levels = disease_order)) %>%
  filter(age %in% c(70, 75, 80, 85, 90)) %>%
  mutate(Algorithm = "Mode approximation")

pdf("../../write/opt_compare.pdf")

resp <- resall %>% 
  dplyr::filter(model %in% c("Independent areas","National"), var=="cf") %>%
  dplyr::filter(area=="Bristol" | (area=="England" & model=="National"), gender=="Female") %>%
  mutate(Algorithm="MCMC") %>%
  full_join(resoptunccf)
resp %>%
  filter(age %in% c(70, 75, 80, 85, 90)) %>%
  mutate(par = 1 - exp(-`50%`),
         plow = 1 - exp(-`2.5%`),
         pupp = 1 - exp(-`97.5%`)) %>%
  ggplot(aes(x=age, y=par, col=algorithm, lty=algorithm)) + 
  geom_errorbar(aes(ymin=plow, ymax=pupp), width=1.5,
              position=position_dodge(width=2), lwd=1.3) +
  geom_point(position=position_dodge(width=2), size=1.3) + 
  facet_wrap(vars(disease), ncol=2, scales="free_y") + 
  xlab("Age (years)") + ylab("Case fatality probability")

dev.off()

## Age curves for all diseases and areas

pdf("../../write/all_cf_indep.pdf",width=6,height=6)

gg <- resall %>% 
  dplyr::filter(model %in% c("Independent areas","National"), var=="cf") %>%
  dplyr::filter(! (disease=="Uterine cancer" & age > 90)) %>%
  mutate(p50 = 1 - exp(-`50%`),
         plow = 1 - exp(-`2.5%`),
         pupp = 1 - exp(-`97.5%`)) %>%
  ggplot(aes(x=age, y=p50, group=area)) + 
  geom_ribbon(aes(ymin=plow, ymax=pupp), 
              fill="deepskyblue") +
  geom_line() + 
  facet_grid(cols=vars(gender), rows=vars(disease), scales = "free_y", switch = "y") + 
  xlab("Age (years)") + 
  ylab("Annual case fatality probability") + 
  xlim(60,100) + 
  theme(strip.text.y.left = element_text(angle = 0))

grob <- ggplotGrob(gg)
# gtable::gtable_show_layout(grob)
idx <- which(grob$layout$name %in% c("panel-5-2","panel-9-2"))
for (i in idx) grob$grobs[[i]] <- grid::nullGrob()
grid::grid.draw(grob)

dev.off()

```

# Hierarchical model

```{r,cache=TRUE}

## UTRC doesn't converge, so we use the empirical Bayes results

resh %>%
  filter(disease=="Stomach cancer", gender=="Male", var=="cf") %>%
  ggplot(aes(x=age, y=`50%`, col=area, group=area)) + 
  geom_line()

resh %>% filter(gender=="Female", disease=="Stomach cancer", 
                var=="cf", age>40, age<60) %>% select(age,`50%`,Rhat) 
resnc <- resh %>% filter(Rhat > 2) 
resnc <- resh %>% filter(Rhat > 1.05) 
table(resnc$disease, resnc$gender)

## Look at precision of CRC estimates as exemplar. 
## Effective sample size of 1000 after a few hours
resh %>% filter(gender=="Male", disease=="Ischemic heart disease", age==60, var=="cf") %>% head
resh %>% filter(gender=="Male", disease=="Colon and rectum cancer", age==60, var=="cf") %>% head

## Between area variability in point and interval ests, hier and nonhier models  
## Specific age and gender, compare between areas 
## Not used in paper

example_age <- 90
dat <- resall %>% filter(age==example_age, var=="cf", 
                         !(disease %in% c("Stomach cancer","Uterine cancer")),
                         #      disease %in% c("Colon and rectum cancer"),
                         model %in% c("Independent areas","Hierarchical"))
ggplot(dat, aes(x=`50%`, y=area, col=model)) +
  #   coord_cartesian(xlim=c(0.01, 0.05)) +
  facet_grid(rows=vars(disease), cols=vars(gender)) + 
  geom_point(position=ggstance::position_dodgev(height=0.4)) +
  geom_errorbarh(aes(xmin=`2.5%`, xmax=`97.5%`),
                 position=ggstance::position_dodgev(height=0.4)) +
  xlab(sprintf("Case fatality for age %s", example_age) )

## Between area variability in age curves, point ests.  CARC.  Shrinkage for oldest
## Use this in paper

pdf("../write/hier_crc.pdf", width=6, height=4)

resall %>% 
  filter(var=="cf", 
         !disease %in% c("Stomach cancer","Uterine cancer"),
         disease %in% c(
           #   "Dementia"
           "Colon and rectum cancer"
           #       "Lung cancer"
         ),
         `97.5%` < 10,
         area %in% cityregions, 
         model %in% c("Independent areas","Hierarchical")) %>%
  mutate(model = relevel(factor(model), "Independent areas")) %>%
  ggplot(aes(x=age, col=area, group=area)) +
  geom_ribbon(aes(ymin=`2.5%`, ymax=`97.5%`), fill="gray80", col="gray80") +
  geom_line(aes(y=`50%`)) + 
  coord_cartesian(ylim=c(0, 1), 
                  xlim=c(70,  100)) +
  scale_x_continuous(breaks=seq(0,100,10)) + 
  ylab("Case fatality rate")  +
  facet_grid(cols=vars(model), rows=vars(gender)) + 
  xlab("Age (years)") + 
  theme_bw() + labs(col="")

dev.off()

```

# Hierarchical model with gender effect independent of area

```{r}

resg <- resg %>% filter(disease != "Stomach cancer") # only national estimates
# convergence check
summary(resg$Rhat)  # 
resnc <- resg %>% filter(Rhat > 2) 
resnc <- resg %>% filter(Rhat > 1.1) 
resnc <- resg %>% filter(Rhat > 1.01) 
table(resnc$disease)
table(resnc$disease, resnc$area)

## point est curves compared between areas 
rescf <- resg %>% filter(gender=="Male", var=="cf")
rescf <- resg %>% filter(gender=="Female", var=="cf")
ggplot(rescf, aes(x=age, col=area)) +
  geom_line(aes(y=`50%`)) + 
  coord_cartesian(ylim=c(0, 1), 
                  xlim=c(50,  100)) +
  scale_x_continuous(breaks=seq(0,100,10)) + 
  ylab("Case fatality (median, 95% CrI)")  +
  facet_wrap(~disease, ncol=2) + 
  xlab("Age (years)")

## Get curves of m/f CF risk ratios by age and area, from the hier model with 
## different gender effect for each area 
reshi <- resh %>%  
  mutate(disease = fct_recode(disease, !!!disease_shorten)) %>%
  filter(! disease %in% c("Breast cancer", "Stomach cancer", "Uterine cancer"), 
         area %in% cityregions, 
         var=="cf") %>%
  mutate(est = log(`50%`), se=(log(`97.5%`)-log(`2.5%`))/4) 
reshie <- reshi %>%
  select(age, gender, area, disease, est) %>% 
  pivot_wider(names_from=gender, values_from=est) %>%
  mutate(lhrmale = Male - Female) %>%
  select(age, area, disease, lhrmale)
reshis <- reshi %>%
  select(age, gender, area, disease, se) %>% 
  pivot_wider(names_from=gender, values_from=se) %>%
  mutate(sediff = sqrt(Male^2 + Female^2)) %>%
  select(age, area, disease, sediff) %>%
  left_join(reshie, by=c("age","area","disease")) %>%
  mutate(est = exp(lhrmale), 
         lower=exp(lhrmale - 2*sediff),
         upper=exp(lhrmale + 2*sediff))

## Superimpose the equivalent common curve from the common effect model
resgi <- resg %>%  
  filter(! disease %in% c("Breast cancer", "Stomach cancer", "Uterine cancer"), 
         area %in% cityregions, 
         var=="cf") %>%
  mutate(disease = fct_recode(disease, !!!disease_shorten)) %>%
  mutate(est = log(`50%`), se=(log(`97.5%`)-log(`2.5%`))/4) 
resgie <- resgi %>%
  select(age, gender, area, disease, est) %>% 
  pivot_wider(names_from=gender, values_from=est) %>%
  mutate(lhrmale = Male - Female) %>%
  select(age, area, disease, lhrmale)
resgis <- resgi %>%
  select(age, gender, area, disease, se) %>% 
  pivot_wider(names_from=gender, values_from=se) %>%
  mutate(sediff = sqrt(Male^2 + Female^2)) %>%
  select(age, area, disease, sediff) %>%
  left_join(resgie, by=c("age","area","disease")) %>%
  mutate(est = exp(lhrmale), 
         lower=exp(lhrmale - 2*sediff),
         upper=exp(lhrmale + 2*sediff))

## Put this in the paper 
## Hazard ratio for men compared to women for case fatality 
## This is good evidence that the general pattern is common among areas 
## TODO Confirm this or otherwise by LOO CV

pdf("../../write/gender_hr.pdf")
ggplot(reshis %>% filter(age > 60), 
       aes(x=age, y=est, group=area, col=area)) + 
  geom_ribbon(aes(ymin=lower, ymax=upper), fill="gray", col=NA, alpha=0.2) + 
  geom_hline(yintercept = 1) +
  geom_line(lwd=0.8) +
  geom_line(data=resgis %>% filter(age>60), col="black", lty=2) +
  facet_wrap(~disease) +
  labs(col="Area") +
  coord_cartesian(ylim=c(0,3)) + 
  ylab("Case fatality rate ratio (men/women)") + xlab("Age (years)")
dev.off()

```

## Lookup table for all results

100 * 9 * 18 * 2 * 3 results for each of inc and CF

TODO just save this data frame in the metahit directory. there's no standard format

TODO append the results from optimisation.  OK to have multiple results for one case 
or just save in different format, since one has mode

```{r}
app_tab <- resall %>%
  filter(var %in% c("cf","inc","rem")) %>%
  select(disease, var, gender, age, area, model, `25%`, `50%`, `97.5%`) %>%
  arrange(disease, var, gender, age, area, model)
app_tab %>% head
```

# Cross validatory comparison

Compare nonhierarchical with hierarchical model, for all diseases and outcomes.

Colon and rectum cancer is highlighted in the paper as a case where the hierarchical model makes a difference to the results, for ages 80+.   The hierarchical model gives better predictions here.

```{r}
looall <- loonh %>% mutate(model="Independent areas") %>%
  full_join(looh %>% mutate(model="Hierarchical")) %>%
  filter(!(model=="National")) %>%
  select(age, gender, elpd_loo, model, area, disease, var) %>%
  group_by(model, gender, disease, var)

looall %>% 
  summarise(eisum = sum(elpd_loo)) %>%
  pivot_wider(names_from=model, values_from=eisum) %>%
  mutate(elpd_diff = `Hierarchical` - `Independent areas`)

looall %>% 
  filter(age > 80) %>%
  summarise(eisum = sum(elpd_loo)) %>%
  pivot_wider(names_from=model, values_from=eisum) %>%
  mutate(elpd_diff = `Hierarchical` - `Independent areas`)%>%
  filter(disease=="Colon and rectum cancer")
```

Graph of how predictive ability depends on age and area, for selected disease and outcome.  Hierarchical model is better for older ages.  Age 50 data seem to be strong enough that shrinkage is not needed.  Note that age 50 parameters are also assumed to apply to younger ages.

```{r}
loosub <- looall %>% 
  filter(disease=="Colon and rectum cancer", gender=="Male", var=="mort") %>%
  ## sum ELPD by five year age group
  pivot_wider(names_from=model, values_from=elpd_loo) %>%
  mutate(elpd_diff = `Hierarchical` - `Independent areas`) %>%
  mutate(agegroup = cut(age, seq(0,100,5), right=FALSE, include.lowest = TRUE)) %>%
  group_by(area, agegroup, gender, var) %>%
  summarise(elpd_diff = sum(elpd_diff), .groups=c("drop")) %>%
  mutate(agemid = as.numeric(as.character(factor(agegroup, labels = seq(2.5, 97.5, by=5)))))

ggplot(loosub, aes(x=agemid, y=elpd_diff, colour=area)) +
  geom_hline(yintercept=0, col="gray", lwd=2) +
  geom_line() + 
  geom_point() + 
  scale_x_continuous(limits=c(30,100), breaks = seq(0,100,10)) + 
  geom_text(data=data.frame(x=c(50,50), y=c(-Inf,Inf)), 
            aes(x=x, y=y), 
            inherit.aes=FALSE,
            label=c("Independent areas preferred","Hierarchical preferred"), 
            vjust=c(0,1), hjust=c(0,0)) +
  directlabels::geom_dl(aes(label=area), 
                        method=list(directlabels::dl.combine("first.points", "last.points")))

```

Then compare hier with hier additive gender

```{r}

## FIXME rerun hier with bugfix in saving loo output .  was triggered when we changed rundf
## to exclude (stmc and?) utrc

looall %>% filter(var=="mort", area=="Bristol", gender=="Male", disease=="Lung cancer", 
                  model=="Hierarchical", age==1)

# elpd : high is good. positive elpd diff favours additive gender model 
loohg <- looall %>% 
  filter(model %in% c("Hierarchical","Hierarchical joint gender")) %>%
  select(var, area, age, gender, disease, model, elpd_loo) %>%
  pivot_wider(names_from=model, values_from=elpd_loo, values_fn=length)

%>%
  mutate(elpd_diff = `Hierarchical joint gender` - `Hierarchical`) %>%
  mutate(agegroup = cut(age, seq(0,100,5), right=FALSE, include.lowest = TRUE)) %>%
  group_by(var, area, gender, disease, agegroup) %>%
  summarise(elpd_diff = sum(elpd_diff), .groups=c("drop")) %>%
  mutate(agemid = as.numeric(as.character(factor(agegroup, labels = seq(2.5, 97.5, by=5)))))

loohg %>% 
  filter(disease %in% c("Ischemic heart disease","Lung cancer")) %>% 
  group_by(agemid, disease) %>%
  summarise(elpd = sum(elpd_diff)) %>%
ggplot(aes(x=agemid, y=elpd, col=disease, group=disease)) + 
  geom_hline(yintercept=0) + geom_line() + geom_point()

# ELPD all over the place for different ages 
# positive: common effects.  negative: different effects 
# Different effects for lung cancer and dem
# not transparent though. 
# IHD vs LC is still biggest contrast 
# not sure makes sense to split by age.  model defines risk as smooth function of age

loohg %>%
  group_by(var, disease) %>%
  summarise(elpd = sum(elpd_diff)) %>%
  arrange(disease) %>%
  pivot_wider(names_from=var, values_from=elpd) %>%
  filter(!is.na(inc)) %>%
  mutate(rem=ifelse(is.na(rem), 0, rem),
         overall = inc+mort+prev+rem)

```